Extended Data

### 2022-03-09 doc/manuscript/Dow_manuscript_Nature.Rmd
a couple small changes requested by R1

To rule out the possibility that observed patterns were strongly influenced by summer drought, we examined the relationship between spring temperatures and summer Standardized Precipitation Evapotranspiration Index[@vicente-serrano_multiscalar_2010].
The latter was obtained from the ForestGEO Climate Data Portal v1.0 (https://forestgeo.github.io/Climate/)[@anderson-teixeira_forestgeo_2020;@begueria_standardized_2014;@vicente-serrano_multiscalar_2010a]. 
Linear models were run with 4-, 6-, and 12-month SPEI values of June, July, and August vs April $T_{max}$ to determine if warm spring temperatures were associated with greater summer drought stress in our data set. 
No significant correlations were found (all p>0.05).

@@

For each species-site combination, we converted tree-ring records into the dimensionless RWI to emphasize interannual variability associated with climate [@cook_methods_1990].
A 2/3rds _n_ spline was applied to each core using ARSTAN to produce standardized ring-width series; _n_ is the number of years in each series[@cook_time_1985;@cook_methods_1990].
An adaptive power transformation, a process that also stabilises the variance over time[@cook_calculating_1997], was used to minimize the influence of outliers in all series. 
Low series replication, often in the earliest portions of a chronology collection, can also inflate the variance of tree-ring records[@jones_estimating_1997].
The 1/3rds spline method was chosen when replication in the inner portion of each chronology (ca. inner 30–50 yr of each record depending on full chronology length) was less than three trees. 
When replication was greater than n = 3 trees, we used the average correlation between raw ring-width series (rbar) method.
The robust biweight mean chronology (RWI) for each species-site combination was calculated from the ring-width indices following variance stabilisation[@cook_time_1985].
We defined chronology start year (Extended Data Table 1) as the year where subsample signal strength (SSS) passed a threshold of SSS = 0.8, or where $\ge$ 80% of the population signal was captured in the chronology.

>>>>

To rule out the possibility that observed patterns were strongly influenced by summer drought, we examined the relationship between spring temperatures and summer Standardized Precipitation Evapotranspiration Index (SPEI)[@vicente-serrano_multiscalar_2010].
The latter was obtained from the ForestGEO Climate Data Portal v1.0 (https://forestgeo.github.io/Climate/)[@anderson-teixeira_forestgeo_2020;@begueria_standardized_2014;@vicente-serrano_multiscalar_2010a]. 
Linear models were run with 4-, 6-, and 12-month SPEI values of June, July, and August vs April $T_{max}$ to determine if warm spring temperatures were associated with greater summer drought stress in our data set. 
No significant correlations were found (all p>0.05).

@@

For each species-site combination, we converted tree-ring records into the dimensionless RWI to emphasize interannual variability associated with climate[@cook_methods_1990].
A 2/3rds _n_ spline was applied to each core using ARSTAN to produce standardized ring-width series; _n_ is the number of years in each series[@cook_time_1985;@cook_methods_1990].
An adaptive power transformation, a process that also stabilises the variance over time[@cook_calculating_1997], was used to minimize the influence of outliers in all series. 
Low series replication, often in the earliest portions of a chronology collection, can also inflate the variance of tree-ring records[@jones_estimating_1997].
The 1/3rds spline method was chosen when replication in the inner portion of each chronology (ca. inner 30–50 yr of each record depending on full chronology length) was less than three trees. 
When replication was greater than n = 3 trees, we used the average correlation between raw ring-width series (rbar) method.
The robust biweight mean chronology (RWI) for each species-site combination was calculated from the ring-width indices following variance stabilisation[@cook_time_1985].
We defined chronology start year (Extended Data Table 1) as the year where subsample signal strength (SSS) passed a threshold of SSS = 0.8, or where $\ge$ 80% of the population signal was captured in the chronology.
